### cleanrmd::html_document_clean

Create a new R Markdown document using the *Clean HTML R Markdown* template in RStudio, or add the following to your `.Rmd` YAML header.

``` yaml
output: 
  cleanrmd::html_document_clean:
    theme: no-class
```

To explore the available themes, set `theme` to `NULL`.

``` yaml
output: 
  cleanrmd::html_document_clean:
    theme: NULL
```

MathJax and FontAwesome are available but disabled by default. To enable either feature, you can set `mathjax` to `local` or `default`, as in `rmarkdown::html_document()`. Set `use_fontawesome` to `TRUE` to enable Font Awesome icons.

``` yaml
output: 
  cleanrmd::html_document_clean:
    mathjax: default
    use_fontawesome: true
```

### Just the theme

You can also load the CSS theme dependencies in other places where [htmltools](https://rstudio.github.io/htmltools/) can be used to provide HTML dependencies, such as Shiny apps.

To include a theme in your app or page, use

```{r eval=FALSE}
cleanrmd::use_cleanrmd(theme = "new.css")
```

To view the list of theme options view the help pages of `?cleanrmd_themes()` or use its output:

```{r}
cleanrmd::cleanrmd_themes()
```
