### cleanrmd::html_document_clean

Create a new R Markdown document using the *Clean HTML R Markdown* template in RStudio, 
or add the following to your `.Rmd` YAML header to use
[cleanrmd::html_document_clean](https://pkg.garrickadenbuie.com/cleanrmd/reference.html#html_document_clean).

``` yaml
output: 
  cleanrmd::html_document_clean:
    theme: no-class
```

To explore the available themes, set `theme` to `NULL`.

``` yaml
output: 
  cleanrmd::html_document_clean:
    theme: NULL
```

```{r include = FALSE}
combine_code <- function(x) {
  knitr::combine_words(x, before = "`")
}
```

**Syntax highlighting** is provided by default by pandoc,
where syntax highlighting is performed during during the render,
minimizing dependencies.
pandoc's highlighting themes include
`r combine_code(cleanrmd:::pandoc_highlight_styles()[-1:-3])`.
The `default` highlighting theme is `arrow`, provided by the [rmarkdown](https://rmarkdown.rstudio.com/docs/) package in addition to the `rstudio` theme.

`html_document_clean()` can also use [Prism](https://prismjs.com/) for highlighting.
In this case, highlighting is performed in the browser
and the dependencies are downloaded as needed.
To use Prism, set `theme: prism` 
or use one of the following value to choose a specific Prism theme:
`r combine_code(cleanrmd:::prism_themes()[-1])`.

**MathJax and FontAwesome** are also available but disabled by default. To enable either feature, you can set `mathjax` to `local` or `default`, as in `rmarkdown::html_document()`. Set `use_fontawesome` to `TRUE` to enable Font Awesome icons.

``` yaml
output: 
  cleanrmd::html_document_clean:
    mathjax: default
    use_fontawesome: true
```

### Just the theme

You can also load the CSS theme dependencies in other places where [htmltools](https://rstudio.github.io/htmltools/) can be used to provide HTML dependencies, such as Shiny apps.

To include a theme in your app or page, use

```{r eval=FALSE}
cleanrmd::use_cleanrmd(theme = "new.css")
```

To view the list of theme options view the help pages of [cleanrmd_themes()](https://pkg.garrickadenbuie.com/cleanrmd/reference.html#cleanrmd_themes) or use its output:

```{r}
cleanrmd::cleanrmd_themes()
```
