---
title: "cleanrmd Reference"
subtitle: "Clean HTML R Markdown Documents"
author: "Garrick Aden-Buie"
date: 'Updated: `r strftime(Sys.time(), "%B %d, %Y")`<br>Version: `r packageVersion("cleanrmd")`'
output: 
  cleanrmd::html_document_clean:
    theme: NULL
    lib_dir: libs
    self_contained: false
    dev: svg
    title_in_header: true
    toc: true
    toc_depth: 2
    pandoc_args: ["--variable", "toc-title=On This Page"]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE)
set.seed(42)
```

```{r metathis, echo = FALSE}
metathis::meta_social(
  title = rmarkdown::metadata$title,
  description = rmarkdown::metadata$subtitle,
  url = "https://pkg.garrickadenbuie.com/cleanrmd/",
  image = "https://pkg.garrickadenbuie.com/cleanrmd/cleanrmd-social.png",
  image_alt = "cleanrmd: shows an array of screenshots of a simple R Markdown report in a variety of styles, fanned out like a hand of cards.",
  twitter_card_type = "summary_large_image",
  twitter_creator = "@grrrck"
)
```

```{r include=FALSE}
library(cleanrmd)
```

```{=html}
<nav>
  <h2>Package Links</h2>
  <ul>
    <li><a href="index.html">Home</a></li>
    <li><a href="https://github.com/gadenbuie/cleanrmd">GitHub</a></li>
    <li><a href="https://cran.r-project.org/web/packages/cleanrmd/index.html">CRAN</a></li>
  </ul>
</nav>
```
## `cleanrmd_themes()` - List cleanrmd themes {#cleanrmd_themes}

### Description

Lists the available themes in [cleanrmd]{.pkg}

### Usage

``` R
cleanrmd_themes()
```

### Value

A character string of available theme names.

### cleanrmd theme list

[cleanrmd]{.pkg} includes the following no-class CSS themes:

- [almond](https://github.com/alvaromontoro/almond.css)

- [awsm.css](https://igoradamenko.github.io/awsm.css/)

- [axist](https://ruanmartinelli.github.io/axist/)

- [bamboo](https://rilwis.github.io/bamboo/)

- [bullframe](https://marcopontili.com/projects/bullframe-css/)

- [holiday](https://holidaycss.js.org/)

- [kacit](https://hakanalpay.com/kacit/)

- [latex.css](https://github.com/vincentdoerig/latex-css)

- [markdown-splendor](http://markdowncss.github.io/splendor/)

- [markdown-retro](http://markdowncss.github.io/retro/)

- [markdown-air](http://markdowncss.github.io/air/)

- [markdown-modest](http://markdowncss.github.io/modest/)

- [marx](https://mblode.github.io/marx/)

- [minicss](https://minicss.us/)

- [new.css](https://newcss.net/)

- [no-class](https://davidpaulsson.github.io/no-class/)

- [picocss](https://picocss.com/)

- [sakura](https://oxal.org/projects/sakura/)

- [sakura-vader](https://oxal.org/projects/sakura/)

- [semantic](https://dimitrinicolas.github.io/semantic.css/)

- [simplecss](https://simplecss.org/)

- [style-sans](https://ungoldman.com/style.css/)

- [style-serif](https://ungoldman.com/style.css/)

- [stylize](https://pkg.garrickadenbuie.com/stylize.css/)

- [superstylin](https://caiogondim.github.io/superstylin/)

- [tacit](https://yegor256.github.io/tacit/)

- [vanilla](https://vanillacss.com/)

- [water](https://kognise.github.io/water.css/)

- [water-dark](https://kognise.github.io/water.css/)

- [writ](https://writ.cmcenroe.me/)

### Examples

```{r}
cleanrmd_themes()
```

## `html_document_clean()` - Clean Rmarkdown HTML Document {#html_document_clean}

### Description

Clean HTML documents with R Markdown.

### Usage

``` R
html_document_clean(
  ...,
  theme = "no-class",
  css = NULL,
  toc = FALSE,
  toc_depth = 3,
  title_in_header = TRUE,
  mathjax = NULL,
  use_fontawesome = FALSE,
  fig_width = 10,
  fig_height = 7,
  fig_retina = 2,
  keep_md = FALSE,
  dev = "png",
  highlight = "default",
  pandoc_args = NULL,
  extra_dependencies = NULL,
  md_extensions = NULL,
  self_contained = !is.null(theme)
)
```

### Arguments

<table>
<colgroup>
<col style="width: 50%" />
<col style="width: 50%" />
</colgroup>
<tbody>
<tr class="odd" style="vertical-align: top;">
<td><code>...</code></td>
<td><p>Additional function arguments to pass to the base R Markdown HTML
output formatter <code>html_document_base</code></p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>theme</code></td>
<td><p>The class-less CSS theme to use, one of
<code>cleanrmd_themes()</code>.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>css</code></td>
<td><p>CSS and/or Sass files to include. Files with an extension of
.sass or .scss are compiled to CSS via <code>sass::sass()</code>. Also,
if <code>theme</code> is a <code>bslib::bs_theme()</code> object, Sass
code may reference the relevant Bootstrap Sass variables, functions,
mixins, etc.</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>toc</code></td>
<td><p><code>TRUE</code> to include a table of contents in the
output</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>toc_depth</code></td>
<td><p>Depth of headers to include in table of contents</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>title_in_header</code></td>
<td><p>If <code>TRUE</code> (default), the title, subtitle, author, and
date are placed in a <code
style="white-space: pre;">⁠&lt;header&gt;⁠</code> tag. This is
semantically correct HTML but some CSS frameworks work better with this
structure than others.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>mathjax</code></td>
<td><p>Include mathjax. The "default" option uses an https URL from a
MathJax CDN. The "local" option uses a local version of MathJax (which
is copied into the output directory). You can pass an alternate URL or
pass <code>NULL</code> to exclude MathJax entirely.</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>use_fontawesome</code></td>
<td><p>Should links to FontAwesome be included in the HTML document's
<code style="white-space: pre;">⁠&lt;head&gt;⁠</code>? Only enable if you
are including FontAwesome icons in your HTML document.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>fig_width</code></td>
<td><p>Default width (in inches) for figures</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>fig_height</code></td>
<td><p>Default height (in inches) for figures</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>fig_retina</code></td>
<td><p>Scaling to perform for retina displays (defaults to 2, which
currently works for all widely used retina displays). Set to
<code>NULL</code> to prevent retina scaling. Note that this will always
be <code>NULL</code> when <code>keep_md</code> is specified (this is
because <code>fig_retina</code> relies on outputting HTML directly into
the markdown document).</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>keep_md</code></td>
<td><p>Keep the markdown file generated by knitting.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>dev</code></td>
<td><p>Graphics device to use for figure output (defaults to
png)</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>highlight</code></td>
<td><p>Syntax highlight engine and style, either a built-in Pandoc
highlighting theme, a theme provided by <span
class="pkg">rmarkdown</span>, or a <a
href="https://prismjs.com/index.html">prismjs</a> theme (see below).
Pass <code>NULL</code> to prevent syntax highlighting.</p>
<p><strong>Pandoc themes.</strong> By default, uses Pandoc's
highlighting style. Pandoc's built-in styles include "tango",
"pygments", "kate", "monochrome", "espresso", "zenburn", "haddock" and
"breezedark".</p>
<p>Two custom styles are also included, "arrow", an accessible color
scheme, and "rstudio", which mimics the default IDE theme.
Alternatively, supply a path to a <code>.theme</code> to use <a
href="https://pandoc.org/MANUAL.html#syntax-highlighting">a custom
Pandoc style</a>. Note that custom themes require Pandoc 2.0+.</p>
<p><strong>Prism themes.</strong> To use the <a
href="https://prismjs.com/index.html">prismjs</a> syntax highlighter,
pass one of "prism", "prism-coy", "prism-dark", "prism-funky",
"prism-okaidia", "prism-solarizedlight", "prism-tomorrow", or
"prism-twilight". To use an alternate Prism theme file, pass the URL or
path to the theme's CSS file prefixed with "prism:", e.g.
<code>prism:my-theme.css</code>. Note that the Prism dependencies are
not embedded into self-contained documents so they will require an
active internet connection to work.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>pandoc_args</code></td>
<td><p>Additional command line options to pass to pandoc</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>extra_dependencies</code></td>
<td><p>Extra dependencies as a list of the <code>html_dependency</code>
class objects typically generated by
<code>htmltools:htmlDependency()</code>.</p></td>
</tr>
<tr class="odd" style="vertical-align: top;">
<td><code>md_extensions</code></td>
<td><p>Markdown extensions to be added or removed from the default
definition of R Markdown. See the <code>rmarkdown_format</code> for
additional details.</p></td>
</tr>
<tr class="even" style="vertical-align: top;">
<td><code>self_contained</code></td>
<td><p>Produce a standalone HTML file with no external dependencies,
using data: URIs to incorporate the contents of linked scripts,
stylesheets, images, and videos. Note that even for self contained
documents MathJax is still loaded externally (this is necessary because
of its size).</p></td>
</tr>
</tbody>
</table>

### Value

An R Markdown output format that can be used with
`⁠output:⁠`{style="white-space: pre;"} in an `.Rmd` or for use with
`rmarkdown::render()`.

### MathJax

Note that MathJax is disabled by default to reduce the overall size of
the final document. You can enable MathJax by setting
`mathjax = "default"`, see `rmarkdown::html_document()` for more
options.

### See Also

`use_cleanrmd()` for using cleanrmd themes in places other than in R
Markdown documents

### Examples

```{r}
if (interactive()) {
  html_document_clean()
}

if (interactive()) {
rmarkdown::render("input.Rmd", html_document_clean())

}
```

## `use_cleanrmd()` - Use a clean CSS theme from cleanrmd {#use_cleanrmd}

### Description

Provides a [cleanrmd]{.pkg} CSS theme using [htmltools]{.pkg}. You can
use this CSS theme anywhere that HTML dependencies are handled via
`htmltools::htmlDependency()`, for example in R Markdown or Quarto
documents or in Shiny apps.

### Usage

``` R
use_cleanrmd(name = NULL)

cleanrmd_theme_dependency(name = NULL)
```

### Arguments

|        |                                                                                                                                                           |
|--------|-----------------------------------------------------------------------------------------------------------------------------------------------------------|
| `name` | The name of the theme, see `cleanrmd_themes()` for a list of available themes. If `NULL`, all themes will be loaded with a simple drop down theme picker. |

### Value

`use_cleanrmd()` returns an `htmltools::tagList()` with an
`htmltools::htmlDependency()`. `cleanrmd_theme_dependency()` returns
only the `htmltools::htmlDependency()`.

### Functions

- `use_cleanrmd()`: Use a clean CSS theme in the current document or app
  (general usage)

- `cleanrmd_theme_dependency()`: Use a clean CSS theme dependency
  (advanced usage)

### R Markdown documents

In [R Markdown](https://rmarkdown.rstudio.com/) (static or Shiny
prerendered), you should use the `html_document_clean()` output format
to use a cleanrmd theme.

::: {.sourceCode .markdown}
    ---
    output:
      cleanrmd::html_document_clean:
        theme: NULL # or pick a specific theme
        self_contained: false
    ---
:::

### Quarto documents

You can also use [cleanrmd]{.pkg} in [Quarto](https://quarto.org/)
documents or apps (using `server: shiny`). You\'ll need to turn off the
themes provided by Quarto with `theme: none` and then call
`cleanrmd::use_cleanrmd()` in a code chunk in your document.

::: {.sourceCode .markdown}
    ---
    title: "Untitled"
    format:
      html:
        theme: none
    #server: shiny
    ---

    ```{r cleanrmd, echo=FALSE}
    cleanrmd::use_cleanrmd("bamboo")
    ```
:::

### Shiny apps

In Shiny apps, you\'ll need to use `shiny::basicPage()` rather than
`shiny::fluidPage()`. Then call `use_cleanrmd()` in your app to use a
cleanrmd theme.

::: {.sourceCode .r}
    library(shiny)

    ui <- shiny::basicPage(
      cleanrmd::use_cleanrmd(),

      h2("Old Faithful Geyser Data"),

      sliderInput(
        "bins",
        "Number of bins:",
        min = 1,
        max = 50,
        value = 30
      ),
      plotOutput("distPlot")
    )
:::

### Examples

```{r}
page <- htmltools::withTags(
  main(
    h2("Small Demo"),
    p("Clean, simple, classy but class-less:"),
    ul(
      li("Works almost anywhere"),
      li("Small and simple"),
      li("Easy to extend"),
      li("Good enough but not perfect")
    )
  )
)

# no styles
if (interactive()) {
  htmltools::browsable(page)
}

# all clean styles
page_clean <- htmltools::tagList(page, use_cleanrmd())
if (interactive()) {
  htmltools::browsable(page_clean)
}

# one clean style
page_water <- htmltools::tagList(page, use_cleanrmd("water"))
if (interactive()) {
  htmltools::browsable(page_water)
}

cleanrmd_theme_dependency("bamboo")
```

